---
title: "envlaster parellelization summary doc"
author: "Sicong He"
date: "10/20/2021"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(eval = FALSE, echo = TRUE)
```
import libraries:
```{r,message=FALSE,warning=FALSE}
library("devtools")
library("roxygen2")
library("testthat")
library("knitr")
library(foreach)
library(doParallel)
library(aster2)
library(envlpaster)
library(doSNOW)
library(aster)
```


`fitness_boot`, `secondboot` in the below section are updated version of the functions based on the 4.2.1 version of R matrix array datatype and could have potential error in the future. Function `fitness_boot` is modified in line 112 compared to the current parallel version. noted that is.matrix() is used in place of class(amat)=="matrix"/"array", similar in line 336 in `secondboot`. Also verify if this change is made in all other version of the function before running the code or it could cause error.

Combined version of paralleled bootstrap function using lapply and foreach in `fit_boot_Efron-combined.R`, does not work as expected

```{r fitboot par function, echo=FALSE}
fitness_boot <- function(model, nboot, index, vectors = NULL, dim = NULL,
    data, amat, newdata, modmat.new = NULL, renewdata = NULL,
    criterion = c("AIC","BIC","LRT"), alpha = 0.05, fit.name = NULL,
    method = c("eigen","1d"), quiet = FALSE,corenum=NULL)
{


  # extract important envelope initial quantities
  # obtain the target and nuisance parameters
  # build U w.r.t. the target parameters
  beta <- model$coef
  x <- model$x
  vars <- colnames(x)
  code <- data$code
  families <- data$families
  fam <- model$fam
  pred <- model$pred
  root <- model$root
  n <- nrow(x)
  nnode <- ncol(x)
  npop <- nrow(newdata) 
  p <- length(beta)
  formula <- model$formula
  modmat.model <- model$modmat
  modmat.mat <- matrix(modmat.model, nrow = n * nnode)
  dimensions <- dim(modmat.model)
  offset <- as.vector(model$origin)

  # obtain tau
  mu <- predict(model, parm.type = "mean.value",
    model.type = "unconditional")
  tau <- crossprod(modmat.mat, mu)
  nuis.ind <- c(1:p)[!index]
  k <- length(index)
  target <- tau[index]

  # obtain the projection used to construct the envelope
  # estimator
  avar <- (model$fisher)[index,index]
  U <- P <- NULL; u <- length(index)
  if(method == "eigen"){
    u <- length(vectors)
    eig <- eigen(avar, symmetric = TRUE)
    G <- eig$vec[,c(vectors)]
    P <- tcrossprod(G)
  }
  if(method == "1d"){
    u <- dim
    U <- target %o% target
    G <- manifold1Dplus(M = avar, U = U, u = u)
    P <- tcrossprod(G)
  }

  tau.env <- crossprod(P,target)
  fulltau <- tau
  fulltau[index] <- tau.env

  # change the model matrix for the envelope estimator
  modelmatrix.int <- modmat.mat
  modelmatrix.int[, index] <- modelmatrix.int[, index] %*% P
  modmat.model.int <- array(modmat.mat, dimensions)

  # convert from tau to beta changed from fulltau to tau
  beta.foo <- transformUnconditional(parm = fulltau,
    modelmatrix.int, data, from = "tau", to = "beta",
    offset = offset)

  # set up for the bootstrap for the MLE
  theta.hat2 <- predict(model, model.type = "cond",
    parm.type = "canon")
  theta.hat2 <- matrix(theta.hat2, nrow = n, ncol = nnode)
  MLE.tau.boot <- matrix(nrow = npop, ncol = nboot[1]) # changed from nrow = p
  b2 <- "try-error"; class(b2) <- "try-error"

  # set up for the bootstrap for the envelope estimator
  # (this envelope estimator corresponds to the selected method)
  theta.hat <- transformUnconditional(parm = fulltau,
    modelmatrix.int, data, from = "tau", to = "theta",
    offset = offset)
  theta.hat <- matrix(theta.hat, nrow = n, ncol = nnode)
  #theta.hat <- theta.hat2
  b <- "try-error"; class(b) <- "try-error"


  # the array (matrix) that specifies Darwinian fitness, noted that is.matrix() is used in place of class(amat)=="matrix"/"array"
  amat.mat <- NULL
  if(is.matrix(amat)==FALSE){
    amat.mat <- matrix(amat, nrow = npop,
    byrow = TRUE)
  }
  if(is.matrix(amat) ==TRUE){
    amat.mat <- amat
  }

  # initial quantities
  est <- est2 <- matrix(0, ncol = npop)
  env.tau.boot <- matrix(0, ncol = p)
  P.list <-  NULL
  class(P.list) <- class(P.list) <- "list"
  vectors.list  <- NULL
  class(vectors.list)  <- "list"
  out <- NULL
  table <- NULL


  #modmat.renew <- model.matrix(m1$formula, data = renewdata)
  cond <- !(colnames(modmat.new) %in% model$dropped)
  modmat.new <- modmat.new[, cond]
  M1.renew <- t(modmat.new[,-index])
  M2.renew <- P %*% t(modmat.new[,index])
  modmat.env.renew <- t(rbind(M1.renew,M2.renew))
  data.renew <- asterdata(newdata, vars = vars, pred = pred,
    group = rep(0, length(vars)), code = code,
    families = families)
  origin.renew <- model$origin[1:npop,]
  offset.renew <- as.vector(origin.renew)

  # fit the aster model using different methods

  check_error = function(class, modmat_status=modmat.model) {
    if(class == "try-error"){
      class <- class(try(aout4star2 <- aster(xstar2, root, pred,
        fam, modmat_status, parm = beta,
        method = "nlm"), silent = TRUE))[1]
    }

    if(class == "try-error"){
      class <- class(try( aout4star2 <- aster(xstar2, root,
        pred, fam, modmat_status), silent = TRUE))[1]
    }

    if(class == "try-error"){
      class<- class(try(aout4star2 <- aster(xstar2, root, pred,
        fam, modmat_status, method = "nlm"), silent = TRUE))[1]
    }
    return(class)
  }
  method_check = function(method, tau, index, model, data, alpha) {
    result = 0
    if (method == "1d") {
      barbaz <- envlpaster::selection(tau, index, model, data,
        alpha = alpha, type = "mean-value", method = "1d")
      if(criterion == "AIC") result <- barbaz$aic
      if(criterion == "BIC") result <- barbaz$bic
      if(criterion == "LRT") result <- barbaz$LRT
    }
    if(method == "eigen"){
      fubar <- envlpaster::selection(tau, index, model, data,
        alpha = alpha, type = "mean-value", method = "eigen")
      if(criterion == "AIC") result <- fubar$aic
      if(criterion == "BIC") result <- fubar$bic
      if(criterion == "LRT") result <- fubar$LRT
    }
    return(result)
  }


  #list of vars and functs that need to be move into the parallel from global
  cl <- makeCluster(corenum*2,type="SOCK")
  registerDoSNOW(cl)
  clusterExport(cl, c('check_error','method_check','theta.hat2',
    'pred', 'fam', 'root','modmat.model','beta',
    'modmat.mat','offset','amat.mat','vectors.list',
    'MLE.tau.boot','est2','b2', 'offset.renew','index','method','data','data.renew','alpha',
    'modmat.new','modmat.model.int','n','nnode','vectors'),
    envir=environment())
  clusterEvalQ(cl,library("aster2","envlpaster","aster"))
  resultlist<-foreach(k=1:nboot[1]) %dopar%{

    # fit the aster model to the resampled data obtained from
    # the MLE of theta
    xstar2 <-aster::raster(theta.hat2, pred, fam, root)
  # fit the aster model to the resampled data obtained from
    # the MLE of theta
    class(b2) <- class(try(aout4star2 <-aster::aster(xstar2, root, pred,
      fam, modmat.model, parm = beta), silent = TRUE))[1]
    class(b2) <- check_error(class(b2))
    # MLE of expected Darwinian fitness
    tau.renew <- transformUnconditional(aout4star2$coef, data,
      modmat = modmat.mat, from = "beta", to = "tau",
      offset = offset)
    phi.MLE.renew <- offset.renew + modmat.new %*% aout4star2$coef
    mu.MLE.renew <- transformSaturated(parm = phi.MLE.renew,
      data.renew, from = "phi", to = "mu")
    MLE.tau.boot <- tau.renew
    est2 <- amat.mat %*% mu.MLE.renew #Dar.fit.MLE
    # selection of the envelope eigenstructure
    vectors.list = method_check(method, tau = tau.renew, index=index, model=aout4star2, data=data, alpha=alpha)
    dim.1d=0
    if (method=="1d") {
      dim.1d=vectors.list
    }

    # get the bootstrapped envelope estimators
    M <- matrix(modmat.model.int, nrow = n * nnode)
    avar.star <- (aout4star2$fisher)[index,index]
    beta.env.star <- aout4star2$coef
    mu.star <- predict(aout4star2, parm.type = "mean.value",
      model.type = "unconditional")
    modelmatrix.star <- M

    tau.env.star <- crossprod(modelmatrix.star, mu.star)
    env.tau.boot <- tau.env.star

    # envelope estimator of expected Darwinian fitness using
    # eigenstructures
    if(method == "eigen"){
      G.star <- eigen(avar.star, symmetric = TRUE)$vec[,c(vectors)]
      P.star <- tcrossprod(G.star)
      P.list <- P.star
      modelmatrix.star[, index] <- modelmatrix.star[, index] %*% P.star
      M1.renew <- t(modmat.new[,-index])
      M2.renew <- P.star %*% t(modmat.new[,index])
      modmat.env.renew <- t(rbind(M1.renew,M2.renew))
    }

    # envelope estimator of expected Darwinian fitness using the
    # 1d algorithm
    if(method == "1d"){
      if(vectors.list < length(index)){
        up.env.star <- crossprod(M, mu.star)[index]
        U.star <- up.env.star %o% up.env.star
        G.star <- manifold1Dplus(avar.star, U = U.star, u = dim.1d)
        P.star <- tcrossprod(G.star)
        P.list[k] <- P.star

        modelmatrix.star[, index] <- modelmatrix.star[, index] %*% P.star
        beta.env.star <- transformUnconditional(parm = tau.env.star,
          modelmatrix.star, data, from = "tau", to = "beta",
          offset = offset)
        modmat.env.renew <- modmat.new

      }
      if(dim.1d == length(index)){
        P.list <- diag(length(index))
        env.tau.boot<- tau.renew
      }
      if(dim.1d == length(index)){
        env.tau.boot<- tau.renew
      }
    }
    # P.list
    # env.tau.boot
    beta.env.star <- transformUnconditional(parm = tau.env.star,
        modelmatrix.star, data, from = "tau", to = "beta",
        offset = offset)
    modmat.env.renew[, index] <- modmat.env.renew[, index] %*% P.star
        mu.env.renew <- transformUnconditional(parm = beta.env.star,
    modmat.env.renew, data.renew, from = "beta", to = "mu",
          offset = offset.renew)

    # the stored expected Darwinian fitness estimates
    est<- amat.mat %*% mu.env.renew
    list(est,
         est2,
         MLE.tau.boot,
        env.tau.boot,
        P.list,
        vectors.list)
  }
  stopImplicitCluster()
  stopCluster(cl)
  env.boot.out=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][1])
  }))
   MLE.boot.out=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][2])
  }))
  MLE.tau.boot=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][3])
  }))
  env.tau.boot=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][4])
  }))
   P.list=do.call('c', lapply(1:nboot[1], function(j){
    resultlist[[j]][5]
  }))
  vectors.list=do.call('c', lapply(1:nboot[1], function(j){
    resultlist[[j]][6]
  }))
  out<-list( env.boot.out = env.boot.out, MLE.boot.out = MLE.boot.out, MLE.tau.boot = MLE.tau.boot,
         env.tau.boot = env.tau.boot,
         P.list = P.list,
         vectors.list = vectors.list)
  return(out)

}
```

```{r secondboot, echo=FALSE}
secondboot <- function(k, nboot2, out, model, index, data, amat, 
    newdata, method = c("eigen","1d"))
{

  # extract necessary components from the top-level of
  # bootstrapping 
  env.boot.out <- out$env.boot.out
  env.1d.boot.out <- out$env.1d.boot.out
  MLE.boot.out <- out$MLE.boot.out
  MLE.tau.boot <- out$MLE.tau.boot
  env.tau.boot <- out$env.tau.boot
  env.1d.tau.boot <- out$env.1d.tau.boot
  P.list <- out$P.list
  P.1d.list <- out$P.1d.list
  vectors.list <- out$vectors.list
  u.1d.list <- out$u.1d.list
  nboot <- ncol(env.boot.out)
  npop <- nrow(env.boot.out)


  # specify necessary quantities for secondboot function
  # not extracted in the above
  aout4star2 <- b2 <- model
  beta <- model$coef
  p <- length(beta)
  n <- nrow(model$x)
  nnode <- ncol(model$x)
  modmat.mat <- matrix(model$modmat, nrow = n * nnode)
  mu <- predict(model, parm.type = "mean.value", 
    model.type = "unconditional")
  #tau <- crossprod(modmat.mat, mu)
  tau <- MLE.tau.boot[, k]
  offset <- as.vector(model$origin)
  code <- data$code
  families <- data$families
  vars <- colnames(model$x)
  fam <- model$fam
  pred <- model$pred
  root <- model$root  


  # necessary for trial run
  cov <- V <- var.Efron <- NULL
  MLE.tau.boot.subsample <- NULL
  est.env.subsample <- NULL


  # the array (matrix) that specifies Darwinian fitness
  amat.mat <- NULL
  if(is.matrix(amat)==FALSE){
    amat.mat <- matrix(amat, nrow = npop, 
    byrow = TRUE)
  }
  if(is.matrix(amat) ==TRUE){
    amat.mat <- amat
    amat <- array(amat.mat, dim = c(npop, nnode, p))
  }

  # initialize important quantities
  P.foo <- M.foo <- fit.foo <- foo <- NULL
  vectors.foo <- u <- NULL

  if(method == "eigen"){
    foo <- env.tau.boot[, k] 
    vectors.foo <- vectors.list[[k]]
    u <- length(vectors.foo)
    P.foo <- P.list[[k]] 
    M.foo <- modmat.mat; M.foo[, index] <- M.foo[, index] %*% P.foo
    fit.foo <- rowSums(out$env.boot.out) / nboot
  }

  if(method == "1d"){
    u <- u.1d.list[[k]]
    fit.foo <- rowSums(out$env.1d.boot.out) / nboot
    if(u == length(index)){
      foo <- MLE.tau.boot[, k]
      P.foo <- P.1d.list[[k]] 
      M.foo <- modmat.mat
    }
    if(u < length(index)){
      foo <- env.1d.tau.boot[, k] 
      P.foo <- P.1d.list[[k]] 
      M.foo <- modmat.mat; M.foo[, index] <- M.foo[, index] %*% P.foo
    }
  }

  # theta values
  theta.samp <- transformUnconditional(foo, M.foo, data, 
    from = "tau", to = "theta", offset = offset)
  theta.samp <- matrix(theta.samp, nrow = n, ncol = nnode)
  
  # setup of model matrices and asterdata object 
  # obtained from top level of bootstrapping
  cond <- !(colnames(modmat.renew) %in% model$dropped)
  modmat.renew <- modmat.env.renew <- modmat.renew[, cond]
  modmat.env.renew[, index] <- 
    modmat.env.renew[, index] %*% P.foo
  data.renew <- asterdata(newdata, vars = vars, pred = pred,
    group = rep(0, length(vars)), code = code, 
    families = families)
  origin.renew <- model$origin[1:npop,]
  offset.renew <- as.vector(origin.renew)    

  # initial quantities
  #M.foo.array <- array(M.foo, dim = dim(model$modmat))
  #print(dim(M.foo.array))
  #print(dim(model$modmat))
  r <- length(index)
  U.2nd <- P.2nd <- Sigma.uu <- matrix(0, r, r)
  Gamma.2nd <- matrix(0, nrow = r, ncol = u)
  M.2nd <- M.foo
  est.env.subsample <- matrix(0, nrow = npop, ncol = nboot2)
  MLE.tau.boot.subsample <- matrix(0, nrow = p, ncol = nboot2)
  sd.Efron <- 0

  # generate many new resamples    
  for(j in 1:nboot2){

    xstar.samp <- raster(theta.samp, pred, fam, root)
    
    # fit the aster model using the secondary generated data
    class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root, 
      pred, fam, model$modmat, parm = beta), silent = TRUE))[1]

    if(class(b2) == "try-error"){
      class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root, pred,
        fam, model$modmat, parm = beta, 
        method = "nlm"), silent = TRUE))[1] 
    }

    if(class(b2) == "try-error"){
      class(b2) <- class(try( aout4star2 <- aster(xstar.samp, root, 
        pred, fam, model$modmat), silent = TRUE))[1]
    }

    if(class(b2) == "try-error"){
      class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root, pred,
        fam, model$modmat, method = "nlm"), silent = TRUE))[1] 
    }


    # get tau and mu from this fit
    mu.star <- predict(aout4star2, parm.type = "mean.value", 
      model.type = "unconditional")

    #tau.renew <- transformUnconditional(aout4star2$coef, data,
    #  modmat = modmat.mat, from = "beta", to = "tau",
    #  offset = offset)      

    # get beta and tau using the model matrix containing
    # the projection into the envelope
    Sigma.uu <- aout4star2$fisher[index, index]
    if(method == "eigen"){
      if(length(vectors.foo) < length(index)){
        Gamma.2nd <- eigen(Sigma.uu, symmetric = TRUE)$vec[, vectors.foo]
        P.2nd <- projection(Gamma.2nd)
        M.2nd[, index] <- M.2nd[, index] %*% P.2nd
      }
      if(length(vectors.foo) == length(index)){
        M.2nd <- M.foo
      }
    }
    if(method == "1d"){
      if(u < length(index)){
        tau.star <- crossprod(M.foo, mu.star)
        U.2nd <- tau.star[index] %*% t(tau.star[index]) 
        Gamma.2nd <- manifold1Dplus(Sigma.uu, U = U.2nd, u = u)
        P.2nd <- tcrossprod(Gamma.2nd)
      }
      if(u == length(index)){ 
        P.2nd <- diag(length(index))
        M.2nd[, index] <- M.foo[, index] %*% P.2nd  
      }
    }


    tau.env.star <- crossprod(M.2nd, mu.star)
    beta.env.star <- transformUnconditional(parm = tau.env.star, 
      M.2nd, data, from = "tau", to = "beta", 
      offset = offset)


    # compute the envelope estimator of expected 
    # Darwinian fitness
    modmat.env.renew[, index] <- modmat.renew[,index] %*% P.2nd
    mu.env.renew <- transformUnconditional(parm = beta.env.star, 
      modmat.env.renew, data.renew, from = "beta", to = "mu", 
      offset = offset.renew)
    est.env.subsample[, j] <- (amat.mat %*% mu.env.renew)
    M.2nd <- M.foo

    # store the canonical statistic value
    MLE.tau.boot.subsample[, j] <- tau.env.star

    # compute the Efron sd estimator99i5
    if(j == nboot2){ 

      B <- t(MLE.tau.boot.subsample - rowSums(MLE.tau.boot.subsample)/nboot2)
      V <- (t(B) %*% B) / nboot2;  eig.V <- eigen(V)
      env.centered <- t(est.env.subsample - fit.foo)
      cov <- t(B) %*% env.centered / nboot2

      var.Efron <- t(cov) %*% eig.V$vec %*% diag(1/eig.V$val) %*% 
        t(eig.V$vec) %*% cov
      sd.Efron <- sqrt(diag(var.Efron))
    }
  }

  ### output 
  out <- list(sd.Efron = sd.Efron, cov = cov, V = V, 
    MLE.tau.boot.subsample = MLE.tau.boot.subsample, 
    est.env.subsample = est.env.subsample)
  return(out)
}
```

Version of combined TWO-LAYER bootstap all paralled using foreach and dosnow clustering method. Error occured as noted `transformUnconditional` around line 1000. This function would result in an error as it reached the second `transformUnconditional`(around line 1000 as in comment). The error would be with large numbers of nboot2(>5), Whereas calling the first layer bootstrap paralleled version(function name  `fitness_boot` in current markdown code) and then the 'secondboot' following the techreport does not generate error.

Combined version of paralleled bootstrap function using lapply and foreach in `fit_boot_Efron-combined.R`, does not work as expected

```{r fitboot par two layer function, echo=FALSE}
fitness_boot2 <- function(model, nboot, index, vectors = NULL, dim = NULL,
    data, amat, newdata, modmat.new = NULL, renewdata = NULL,
    criterion = c("AIC","BIC","LRT"), alpha = 0.05, fit.name = NULL,
    method = c("eigen","1d"), quiet = FALSE,corenum=NULL)
{

  # stopping condition
  #stopifnot(!(is.null(vectors) & is.null(u)))
  #stopifnot(!(is.null(fit.name) & is.null(renewdata)))
  #timer <- proc.time()

  # extract important envelope initial quantities
  # obtain the target and nuisance parameters
  # build U w.r.t. the target parameters
  beta <- model$coef
  x <- model$x
  vars <- colnames(x)
  code <- data$code
  families <- data$families
  fam <- model$fam
  pred <- model$pred
  root <- model$root
  n <- nrow(x)
  nnode <- ncol(x)
  npop <- nrow(newdata) #same as nrow(env.boot.out)?
  p <- length(beta)
  formula <- model$formula
  modmat.model <- model$modmat
  modmat.mat <- matrix(modmat.model, nrow = n * nnode)
  dimensions <- dim(modmat.model)
  offset <- as.vector(model$origin)

  # obtain tau
  mu <- predict(model, parm.type = "mean.value",
    model.type = "unconditional")
  tau <- crossprod(modmat.mat, mu)
  nuis.ind <- c(1:p)[!index]
  k <- length(index)
  target <- tau[index]

  # obtain the projection used to construct the envelope
  # estimator
  avar <- (model$fisher)[index,index]
  U <- P <- NULL; u <- length(index)
  if(method == "eigen"){
    u <- length(vectors)
    eig <- eigen(avar, symmetric = TRUE)
    G <- eig$vec[,c(vectors)]
    P <- tcrossprod(G)
  }
  if(method == "1d"){
    u <- dim
    U <- target %o% target
    G <- manifold1Dplus(M = avar, U = U, u = u)
    P <- tcrossprod(G)
  }

  tau.env <- crossprod(P,target)
  fulltau <- tau
  fulltau[index] <- tau.env

  # change the model matrix for the envelope estimator
  modelmatrix.int <- modmat.mat
  modelmatrix.int[, index] <- modelmatrix.int[, index] %*% P
  modmat.model.int <- array(modmat.mat, dimensions)

  # convert from tau to beta changed from fulltau to tau
  beta.foo <- transformUnconditional(parm = fulltau,
    modelmatrix.int, data, from = "tau", to = "beta",
    offset = offset)

  # set up for the bootstrap for the MLE
  theta.hat2 <- predict(model, model.type = "cond",
    parm.type = "canon")
  theta.hat2 <- matrix(theta.hat2, nrow = n, ncol = nnode)
  MLE.tau.boot <- matrix(nrow = npop, ncol = nboot[1]) # changed from nrow = p
  b2 <- "try-error"; class(b2) <- "try-error"

  # set up for the bootstrap for the envelope estimator
  # (this envelope estimator corresponds to the selected method)
  theta.hat <- transformUnconditional(parm = fulltau,
    modelmatrix.int, data, from = "tau", to = "theta",
    offset = offset)
  theta.hat <- matrix(theta.hat, nrow = n, ncol = nnode)
  #theta.hat <- theta.hat2
  b <- "try-error"; class(b) <- "try-error"


  # the array (matrix) that specifies Darwinian fitness
  amat.mat <- NULL
  if(is.matrix(amat)==TRUE){
    amat.mat <- matrix(amat, nrow = npop,
    byrow = TRUE)
  }
  if(is.matrix(amat)==FALSE){
    amat.mat <- amat
    #amat <- array(amat.mat, dim = c(npop, nnode, p))
  }


  # initial quantities
  est <- est2 <- matrix(0, ncol = npop)
  env.tau.boot <- matrix(0, ncol = p)
  P.list <-  NULL
  class(P.list) <- class(P.list) <- "list"
  vectors.list  <- NULL
  class(vectors.list)  <- "list"
  out <- NULL
  table <- NULL


  #modmat.renew <- model.matrix(m1$formula, data = renewdata)
  cond <- !(colnames(modmat.new) %in% model$dropped)
  modmat.new <- modmat.new[, cond]
  M1.renew <- t(modmat.new[,-index])
  M2.renew <- P %*% t(modmat.new[,index])
  modmat.env.renew <- t(rbind(M1.renew,M2.renew))
  data.renew <- asterdata(newdata, vars = vars, pred = pred,
    group = rep(0, length(vars)), code = code,
    families = families)
  origin.renew <- model$origin[1:npop,]
  offset.renew <- as.vector(origin.renew)

  # fit the aster model using different methods

  check_error = function(class, modmat_status=modmat.model) {
    if(class == "try-error"){
      class <- class(try(aout4star2 <- aster(xstar2, root, pred,
        fam, modmat_status, parm = beta,
        method = "nlm"), silent = TRUE))[1]
    }

    if(class == "try-error"){
      class <- class(try( aout4star2 <- aster(xstar2, root,
        pred, fam, modmat_status), silent = TRUE))[1]
    }

    if(class == "try-error"){
      class<- class(try(aout4star2 <- aster(xstar2, root, pred,
        fam, modmat_status, method = "nlm"), silent = TRUE))[1]
    }
    return(class)
  }
  method_check = function(method, tau, index, model, data, alpha) {
    result = 0
    if (method == "1d") {
      barbaz <- envlpaster::selection(tau, index, model, data,
        alpha = alpha, type = "mean-value", method = "1d")
      if(criterion == "AIC") result <- barbaz$aic
      if(criterion == "BIC") result <- barbaz$bic
      if(criterion == "LRT") result <- barbaz$LRT
    }
    if(method == "eigen"){
      fubar <- envlpaster::selection(tau, index, model, data,
        alpha = alpha, type = "mean-value", method = "eigen")
      if(criterion == "AIC") result <- fubar$aic
      if(criterion == "BIC") result <- fubar$bic
      if(criterion == "LRT") result <- fubar$LRT
    }
    return(result)
  }


  #list of vars and functs that need to be move into the parallel from global
  cl <- makeCluster(corenum*2,type="SOCK")
  registerDoSNOW(cl)
  clusterExport(cl, c('check_error','method_check','theta.hat2',
    'pred', 'fam', 'root','modmat.model','beta',
    'modmat.mat','offset','amat.mat','vectors.list',
    'MLE.tau.boot','est2','b2', 'offset.renew','index','method','data','data.renew','alpha',
    'modmat.new','modmat.model.int','n','nnode','vectors'),
    envir=environment())
  clusterEvalQ(cl,library("aster2","envlpaster","aster"))
  resultlist<-foreach(k=1:nboot[1]) %dopar%{

    # fit the aster model to the resampled data obtained from
    # the MLE of theta
    xstar2 <-aster::raster(theta.hat2, pred, fam, root)
  # fit the aster model to the resampled data obtained from
    # the MLE of theta
    class(b2) <- class(try(aout4star2 <-aster::aster(xstar2, root, pred,
      fam, modmat.model, parm = beta), silent = TRUE))[1]
    class(b2) <- check_error(class(b2))
    # MLE of expected Darwinian fitness
    tau.renew <- transformUnconditional(aout4star2$coef, data,
      modmat = modmat.mat, from = "beta", to = "tau",
      offset = offset)
    phi.MLE.renew <- offset.renew + modmat.new %*% aout4star2$coef
    mu.MLE.renew <- transformSaturated(parm = phi.MLE.renew,
      data.renew, from = "phi", to = "mu")
    MLE.tau.boot <- tau.renew
    est2 <- amat.mat %*% mu.MLE.renew #Dar.fit.MLE
    # selection of the envelope eigenstructure
    vectors.list = method_check(method, tau = tau.renew, index=index, model=aout4star2, data=data, alpha=alpha)
    dim.1d=0
    if (method=="1d") {
      dim.1d=vectors.list
    }

    # get the bootstrapped envelope estimators
    M <- matrix(modmat.model.int, nrow = n * nnode)
    avar.star <- (aout4star2$fisher)[index,index]
    beta.env.star <- aout4star2$coef
    mu.star <- predict(aout4star2, parm.type = "mean.value",
      model.type = "unconditional")
    modelmatrix.star <- M

    tau.env.star <- crossprod(modelmatrix.star, mu.star)
    env.tau.boot <- tau.env.star

    # envelope estimator of expected Darwinian fitness using
    # eigenstructures
    if(method == "eigen"){
      G.star <- eigen(avar.star, symmetric = TRUE)$vec[,c(vectors)]
      P.star <- tcrossprod(G.star)
      P.list <- P.star
      modelmatrix.star[, index] <- modelmatrix.star[, index] %*% P.star
      M1.renew <- t(modmat.new[,-index])
      M2.renew <- P.star %*% t(modmat.new[,index])
      modmat.env.renew <- t(rbind(M1.renew,M2.renew))
    }

    # envelope estimator of expected Darwinian fitness using the
    # 1d algorithm
    if(method == "1d"){
      if(vectors.list < length(index)){
        up.env.star <- crossprod(M, mu.star)[index]
        U.star <- up.env.star %o% up.env.star
        G.star <- manifold1Dplus(avar.star, U = U.star, u = dim.1d)
        P.star <- tcrossprod(G.star)
        P.list[k] <- P.star

        modelmatrix.star[, index] <- modelmatrix.star[, index] %*% P.star
        beta.env.star <- transformUnconditional(parm = tau.env.star,
          modelmatrix.star, data, from = "tau", to = "beta",
          offset = offset)
        modmat.env.renew <- modmat.new

      }
      if(dim.1d == length(index)){
        P.list <- diag(length(index))
        env.tau.boot<- tau.renew
      }
      if(dim.1d == length(index)){
        env.tau.boot<- tau.renew
      }
    }
    # P.list
    # env.tau.boot
    beta.env.star <- transformUnconditional(parm = tau.env.star,
        modelmatrix.star, data, from = "tau", to = "beta",
        offset = offset)
    modmat.env.renew[, index] <- modmat.env.renew[, index] %*% P.star
        mu.env.renew <- transformUnconditional(parm = beta.env.star,
    modmat.env.renew, data.renew, from = "beta", to = "mu",
          offset = offset.renew)

    # the stored expected Darwinian fitness estimates
    est<- amat.mat %*% mu.env.renew
    list(est,
         est2,
         MLE.tau.boot,
        env.tau.boot,
        P.list,
        vectors.list)
  }
  stopImplicitCluster()
  stopCluster(cl)
  env.boot.out=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][1])
  }))
   MLE.boot.out=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][2])
  }))
  MLE.tau.boot=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][3])
  }))
  env.tau.boot=do.call(cbind, lapply(1:nboot[1], function(j){
    unlist(resultlist[[j]][4])
  }))
   P.list=do.call('c', lapply(1:nboot[1], function(j){
    resultlist[[j]][5]
  }))
  vectors.list=do.call('c', lapply(1:nboot[1], function(j){
    resultlist[[j]][6]
  }))
  # out<-list( env.boot.out = env.boot.out, MLE.boot.out = MLE.boot.out, MLE.tau.boot = MLE.tau.boot,
  #       env.tau.boot = env.tau.boot,
  #       P.list = P.list,
  #       vectors.list = vectors.list)
  cl<- NULL

  #secondboot
  ##########################################################################
  ##########################################################################
  ##########################################################################

  # necessary components from the top-level of
  # bootstrapping
  # all outout from 1st boot strap
 

  # specify necessary quantities for secondboot function
  # variables decleared above:
  # beta, model(auto4star2), n, p, nnode, modmat.mat, mu
  # offset, code, familes, vars, fam, pred, root

  # necessary for trial run
  cov <- V <- var.Efron <- NULL
  MLE.tau.boot.subsample <- NULL
  est.env.subsample <- NULL

  # the array (matrix) that specifies Darwinian fitness
  amat.mat <- NULL
  if(is.matrix(amat)==FALSE){
    amat.mat <- matrix(amat, nrow = npop,
    byrow = TRUE)
  }
  if(is.matrix(amat)==TRUE){
    amat.mat <- amat
    #amat <- array(amat.mat, dim = c(npop2, nnode, p))
  }
  # setup of model matrices and asterdata object
  # obtained from top level of bootstrapping
  cond <- !(colnames(modmat.new) %in% model$dropped)
  modmat.new <- modmat.env.renew <- modmat.new[, cond]
  
  data.renew <- asterdata(newdata, vars = vars, pred = pred,
                          group = rep(0, length(vars)), code = code,
                          families = families)
  origin.renew <- model$origin[1:npop,]
  offset.renew <- as.vector(origin.renew)

  #initialize quantities independent of k(RESULT LIST)
  r <- length(index)
  U.2nd <- P.2nd <- Sigma.uu <- matrix(0, r, r)
  aout4star2 <- b2 <- model
  #est.env.subsample <- matrix(0, nrow = npop2, ncol = nboot[2])
  #MLE.tau.boot.subsample <- matrix(0, nrow = p, ncol = nboot[2])
  sd.Efron <- 0

  ######################################################################
  #end of values independent of k, the iter number
  
  cl <- makeCluster(corenum*2,type="SOCK")
  registerDoSNOW(cl)
  clusterExport(cl, c('raster','projection',
    'pred', 'fam', 'root','beta',
    'model','aster','r','vectors.list','modmat.mat','modmat.env.renew','modmat.new', 'MLE.tau.boot','P.list','env.boot.out',
    'P.2nd','U.2nd','data','offset','offset.renew',
    'data.renew','index','amat.mat','nboot','npop','p','method','env.tau.boot','n','nnode','b2','aout4star2'),
    envir=environment())
  clusterEvalQ(cl,library("aster2","envlpaster","aster"))
  resultlist2 <- foreach(k=1:nboot[1]) %dopar%{
    # initialize important quantities
    #involving k(iteration number)
    P.foo <- M.foo <- fit.foo <- foo <- NULL
    vectors.foo <- u <- NULL

    if(method == "eigen"){
      foo <- env.tau.boot[, k]
      vectors.foo <- vectors.list[[k]]
      u <- length(vectors.foo)
      P.foo <- P.list[[k]]
      M.foo <- modmat.mat; M.foo[, index] <- M.foo[, index] %*% P.foo
      fit.foo <- rowSums(env.boot.out) / nboot[1]
    }

    if(method == "1d"){
      u <- vectors.list[[k]]
      fit.foo <- rowSums(env.boot.out) / nboot[1]
      if(u == length(index)){
        foo <- MLE.tau.boot[, k]
        P.foo <- P.list[[k]]
        M.foo <- modmat.mat
      }
      if(u < length(index)){
        foo <- env.tau.boot[, k]
        P.foo <- P.list[[k]]
        M.foo <- modmat.mat; M.foo[, index] <- M.foo[, index] %*% P.foo
      }
    }
    modmat.env.renew[, index] <-modmat.env.renew[, index] %*% P.foo

    # theta values
    theta.samp <- transformUnconditional(foo, M.foo, data,
                                        from = "tau", to = "theta", offset = offset)
    theta.samp <- matrix(theta.samp, nrow = n, ncol = nnode)

    # initial quantities dependent of k
    M.2nd <- M.foo#check if could be simplified
    Gamma.2nd <- matrix(0, nrow = r, ncol = u)
    
    #est.env.subsample<- matrix(0, nrow = npop2, ncol = nboot[2])
    #MLE.tau.boot.subsample <- matrix(0, nrow = p, ncol = nboot[2])
    rlist = NULL
    mle.tao.list = NULL
    #Generate new samples
    for(j in 1:nboot[2]){

      xstar.samp <- raster(theta.samp, pred, fam, root)

      # fit the aster model using the secondary generated data
      class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root,
                                                pred, fam, model$modmat, parm = beta), silent = TRUE))[1]

      if(class(b2) == "try-error"){
        class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root, pred,
                                                  fam, model$modmat, parm = beta,
                                                  method = "nlm"), silent = TRUE))[1]
      }

      if(class(b2) == "try-error"){
        class(b2) <- class(try( aout4star2 <- aster(xstar.samp, root,
                                                    pred, fam, model$modmat), silent = TRUE))[1]
      }

      if(class(b2) == "try-error"){
        class(b2) <- class(try(aout4star2 <- aster(xstar.samp, root, pred,
                                                  fam, model$modmat, method = "nlm"), silent = TRUE))[1]
      }


      # get tau and mu from this fit
      mu.star <- predict(aout4star2, parm.type = "mean.value",
                        model.type = "unconditional")

      # get beta and tau using the model matrix containing
      # the projection into the envelope
      Sigma.uu <- aout4star2$fisher[index, index]
      if(method == "eigen"){
        if(length(vectors.foo) < length(index)){
          Gamma.2nd <- eigen(Sigma.uu, symmetric = TRUE)$vec[, vectors.foo]
          P.2nd <- projection(Gamma.2nd)
          M.2nd[, index] <- M.2nd[, index] %*% P.2nd
        }
        if(length(vectors.foo) == length(index)){
          M.2nd <- M.foo
        }
      }
      if(method == "1d"){
        if(u < length(index)){
          tau.star <- crossprod(M.foo, mu.star)
          U.2nd <- tau.star[index] %*% t(tau.star[index])
          Gamma.2nd <- manifold1Dplus(Sigma.uu, U = U.2nd, u = u)
          P.2nd <- tcrossprod(Gamma.2nd)
        }
        if(u == length(index)){
          P.2nd <- diag(length(index))
          M.2nd[, index] <- M.foo[, index] %*% P.2nd
        }
      }


      tau.env.star <- crossprod(M.2nd, mu.star)
      beta.env.star <- transformUnconditional(parm = tau.env.star,
                                              M.2nd, data, from = "tau", to = "beta",
                                              offset = offset)#!!!
      
    #check
      # compute the envelope estimator of expected
      # Darwinian fitness
      modmat.env.renew[, index] <- modmat.new[,index] %*% P.2nd
      print("parm=")
      print(beta.env.star)
      print("modmat=")
      print(modmat.env.renew)
      print('data')
      print(data.renew)
      print("offset")
      print(offset.renew)
      mu.env.renew <- transformUnconditional(parm = beta.env.star,
                                            modmat.env.renew,#M.2nd#, data.renew, from = "beta", to = "mu",
                                            offset = offset.renew) #!!
      
      #appeared infinite when using nboot=c(10,10) error:Error in { : task 10 failed - "is.finite(theta) are not all TRUE"
      rlist = c(rlist,list(beta.env.star,modmat.env.renew))
      mle.tao.list = cbind(mle.tao.list,tau.env.star)
      # compute the Efron sd estimator99i5
      
    }
    list(rlist,mle.tao.list,fit.foo)
    #list(est.env.subsample =est.env.subsample, MLE.tau.boot.subsample = tau.env.star, fit.foo)
  }
  stopImplicitCluster()
  stopCluster(cl)

  return(resultlist2)
}
```




Setup code for testing the bootstrap
```{r read data}
data(Mguttatus)
```


```{r read vars}
data <- Mguttatus
redata <- Mguttatus.redata
vars <- quantities$vars
pred <- quantities$pred
group <- quantities$group
code <- quantities$code
fam <- quantities$fam
nnode <- length(vars)
n <- nrow(redata) / nnode
families <- quantities$families
root <- redata$root
fit <- redata$fit
varvar <- quantities$varvar
idvar <- quantities$idvar
```


```{r load file}
load("main-all-terms-DavidLowrydata.RData")

```



```{r setup astermodel}
m.main <- aster(resp ~ varb + fit:(gen_bac + site +
inversion + type), pred, fam, varvar = varvar,
idvar = idvar, data = redata, root = root)

```
```{r setup2}
target <- which(grepl("fit", names(m.main$coef)))
avar.targ <- m.main$fisher[target, target]
modmat.mat <- matrix(m.main$modmat, nrow = n*nnode)
tau <- crossprod(modmat.mat, redata$resp)
tau.targ <- tau[target]
beta <- m.main$coef
p <- length(beta)
```

```{r setup3}
foo <- selection(tau, target, m.main, data, type = "mean-value",method = "eigen", alpha = 0.01)
```


```{r setup4}
a <- levels(test$gen_bac)
b <- levels(test$site)
c <- levels(test$inversion)
d <- levels(test$type)
fred <- expand.grid(a = a, b = b, c = c, d = d)
colnames(fred) <- c("gen_bac","site","inversion","type")
fred$sur_flw <- 1
fred$flws <- 1
renewdata <- reshape(fred, varying = list(vars),
                     direction = "long", timevar = "varb",
                     times = as.factor(vars), v.names = "resp")
fit.renew <- as.numeric(grepl("flws", renewdata$varb))
renewdata$fit <- fit.renew
renewdata$root <- 1
npop <- nrow(fred)
modmat.renew <- model.matrix(m.main$formula, data = renewdata)
modmat.renew[, 2] <- 1 - modmat.renew[, 2]
index <- !(colnames(modmat.renew) %in% m.main$dropped)
modmat.renew <- modmat.renew[, index]
amat <- matrix(0, nrow = npop, ncol = nnode * npop)
```


```{r setup5}
for(j in 1:npop){ 
  amat[j, npop + j] =amat[j, npop + j]
}
```


Fitting the model, change`fitness_boot` to `fitness_boot2` would change calculation to two-layer parelled bootstap.
```{r test out,warning=FALSE}
m3 <- fitness_boot(model = m.main, nboot = c(100,10), index = target,
                     vectors = foo$bic, data = data, amat = amat, newdata = fred,
                     modmat.new = modmat.renew, renewdata = renewdata,
                     criterion = "BIC", method = "eigen", quiet = FALSE,corenum =4)
```

Attempt to take out the result from 1st bootstrap(result from `fitness_bot`) and manually tried second bootstrap for output
```{r, eval=FALSE}
   env.sum<- matrix(0,npop, nboot[2])
  sd.Efron.mean <- rep(0, npop)

  for (i in 1:nboot[1]){
    resultlist2[[i]][[1]]=matrix(resultlist2[[i]][[1]], ncol = nboot[2])
    resultlist2[[i]][[1]] =amat.mat%*%resultlist2[[i]][[1]] 
    env.sum = env.sum+resultlist2[[i]][[1]]
    
    B <- t(resultlist2[[i]][[2]] - rowSums(resultlist2[[i]][[2]])/nboot[2])
    V <- (t(B) %*% B) / nboot[2];  eig.V <- eigen(V)
    env.centered <- t(resultlist2[[i]][[1]] - resultlist2[[i]][[3]])
    cov <- t(B) %*% env.centered / nboot[2]
    var.Efron <- t(cov) %*% eig.V$vec %*% diag(1/eig.V$val) %*%
      t(eig.V$vec) %*% cov
    sd.Efron <- sqrt(diag(var.Efron))
    
    sd.Efron.mean = sd.Efron.mean + sd.Efron/nboot[1]
  }
  mu.renew <- transformUnconditional(beta, data.renew,
                                     modmat = modmat.renew, from = "beta", to = "mu",
                                     offset = offset.renew) #from the second bootstrap, using npop2
  fit <- amat%*% mu.renew
  dev<-MLE.boot.out-as.vector(fit)
  mle.sum = (dev[, i] %*% t(dev[, i]))
  for (i in 2:nboot[1]){
    mle.sum <- mle.sum + (dev[, i] %*% t(dev[, i]))
  }
  sd.MLE <- sqrt(diag(mle.sum) / (nboot[1] - 1))
  env.sum = env.sum/nboot[1]
  ratio <- sd.MLE / sd.Efron.mean
  final_list = list(env.sum,sd.Efron.mean,fit,mle.sum,ratio, resultlist2,env.boot.out)
```







